Molecular transport in human pial perivasculature

### 2024-11-27 paper/supplementary_main.tex
Wrap up in Supplementary Methods for now.
Former-commit-id: bc37098f781cc5c6edb3804605a3dc8712f1bbb6

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure \discuss{mass conservation}, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) ni(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n^i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n^i(y_{\rm in}) = 1, \quad
  n^i(y_{\rm{out}}) = -1, \quad
  n^i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
  \label{eq:def:sign}
\end{equation}

Finally, we assume no molecular efflux at all end nodes $\mathcal{T}$
via a homogeneous Neumann condition augmenting~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
    \label{eq:1d_bcs}
\end{equation}

@@

To model steady flow induced by pressure differences between end nodes
in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
internal nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider
the following system of hydraulic network
equations~\cite{tithof2022network, daversin2022geometrically,
  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
\rightarrow \R$, which represent the fluid flux and the average
pressure across cross-sections of the PVSs, respectively. These are
defined for each $\Lambda^i$ by solving:
\begin{subequations}
\begin{alignat}{2}
  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
  \quad \text{ in } \Lambda^i , \\ 
  - \partial_s \hat{q} & = 0 &&
  \quad \text{ in } \Lambda^i ,  
\end{alignat} 
\label{eq:1d_flow}%
\end{subequations}%
where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
The PVS flux $\hat{q}$ and the associated average PVS velocity in the
axial direction $\hat{u}$ are directly related by the cross-sectional area
\begin{equation}
  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
\end{equation}

In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
the fluid pressure and conservation of flux at each internal node. Write
$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
\begin{subequations}
\begin{align}
  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
\end{align}
\end{subequations}
where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
network end nodes:
\begin{equation}
  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{T} .
\end{equation}
In the case where modelling flow in the perivascular networks
$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
fluid pressures $p_0$.

@@

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h^k = \{c \in L^2(\Omega): \;  c \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathbb{P}^k(E)$ denotes the space of
polynomials of degree $\leqslant k$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. \discuss{Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.}
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D \discuss{\phi} \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$, \discuss{$h_F$ denotes the facet diameter},
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = \phi D|_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
\discuss{porosity-weighted} diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h^k$ such that for all $v \in V_h^k$:
\begin{equation}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \discuss{\cup \Gamma_{\rm LV}} } \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm \discuss{S}SAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

To discretize~\eqref{eq:1d_pde} with the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
linear polynomials defined over $\Lambda$:
\begin{equation}
  \hat V_h
  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathbb{P}^1(\Lambda^i) \}, 
\end{equation}
where $\mathbb{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.

@@

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_h &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathbb{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
\mer{MER: How about the homogenized version of $\bm V_h$? - to distinguish for test space?} Here, $H(\mathrm{div}, \Omega)$ is the space of $L^2(\Omega)$ vector fields with $L^2(\Omega)$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. \mer{For
convenience, we collect all facets exterior to the CSF space but not
on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$.} Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
  - \sum_{F \in \mer{\mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}} \left (
  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where \discuss{$h$ denotes the element size} and on exterior facets \mer{the average and jump operators take the one-sided values}. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_h \times Q_h$ such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0, \qquad \foralls \bm v_h \in \bm V_h, \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0, \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

\subsubsection{PVS flow}

To enforce conservation of flux and continuity of pressure on bifurcation points, we approximate the velocity with a piecewise discontinuous polynomials and the pressure with continuous piecewise polynomials, respectively. The finite element spaces thus are:

\begin{align}
   V_h^{\rm PVS} = \{v_h \in L^2(\Lambda): v_h \vert_{\Lambda_i} \in
   \mathbb{P}^{k -1}(\Lambda_i)\} \text{ and }
   Q_h^{\rm PVS} = \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathbb{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}

The weak form reads: Find $(\hat{u}, \hat{p}) \in V_h^{\rm PVS} \times Q_h^{\rm PVS}$ such that

\begin{align}
(A \hat{u}, v) + \left(v, \frac{A \kappa}{\mu} \partial_s \hat{p}  \right) &= 0 \quad \forall v \in V_h^{\rm PVS} \\
(A \hat{u}, \partial_s q ) &= 0 \quad  \forall q \in  Q_h^{\rm PVS}
\end{align}

where $\tau$ is the tangent vector of the network. Here we employ the lowest order discretization and set $k=1$. For discussion of stability of the discrete model we refer to \cite{gjerde2024directional}.

@@

>>>>

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
\hat{c}|_{\Lambda^i}$, and let $\mathcal{E}(y)$ denote the set of
(one, two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
impose continuity of the concentrations, we set that the
concentrations when viewed from each edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure \discuss{mass conservation}, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) ni(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n^i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n^i(y_{\rm in}) = 1, \quad
  n^i(y_{\rm{out}}) = -1, \quad
  n^i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
  \label{eq:def:sign}
\end{equation}

Finally, we assume no molecular efflux at all end nodes $\mathcal{N}$
via a homogeneous Neumann condition augmenting~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
    \label{eq:1d_bcs}
\end{equation}

@@

To model steady flow induced by pressure differences between end nodes
in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
internal nodes $\mathcal{B}$ and end nodes $\mathcal{N}$, we consider
the following system of hydraulic network
equations~\cite{tithof2022network, daversin2022geometrically,
  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
\rightarrow \R$, which represent the fluid flux and the average
pressure across cross-sections of the PVSs, respectively. These are
defined for each $\Lambda^i$ by solving:
\begin{subequations}
\begin{alignat}{2}
  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
  \quad \text{ in } \Lambda^i , \\ 
  - \partial_s \hat{q} & = 0 &&
  \quad \text{ in } \Lambda^i ,  
\end{alignat} 
\label{eq:1d_flow}%
\end{subequations}%
where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
The PVS flux $\hat{q}$ and the associated average PVS velocity in the
axial direction $\hat{u}$ are directly related by the cross-sectional area
\begin{equation}
  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
\end{equation}

In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
the fluid pressure and conservation of flux at each internal node. Write
$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
\begin{subequations}
\begin{align}
  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
\end{align}
\label{eq:1d_flow:bifurcations}
\end{subequations}
where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
network end nodes:
\begin{equation}
  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{N} .
  \label{eq:1d_flow:bcs}
\end{equation}
In the case where modelling flow in the perivascular networks
$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
fluid pressures $p_0$.

@@

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h^k = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^k(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^k(E)$ denotes the space of
polynomials of degree $\leqslant k$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. \discuss{Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.}
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D \discuss{\phi} \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$, \discuss{$h_F$ denotes the facet diameter},
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = \phi D|_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
\discuss{porosity-weighted} diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h^k$ such that for all $v \in V_h^k$:
\begin{equation}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \discuss{\cup \Gamma_{\rm LV}} } \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm \discuss{S}SAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

To discretize~\eqref{eq:1d_pde} with the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
linear polynomials defined over $\Lambda$:
\begin{equation}
  \hat V_h
  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
\end{equation}
where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$.

@@

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_h &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
\mer{MER: How about the homogenized version of $\bm V_h$? - to distinguish for test space?} Here, $H(\mathrm{div}, \Omega)$ is the space of $L^2(\Omega)$ vector fields with $L^2(\Omega)$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_h \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. \mer{For
convenience, we collect all facets exterior to the CSF space but not
on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$.} Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
  - \sum_{F \in \mer{\mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}}} \left (
  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \frac{\sigma \mu}{h} \int_{F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where \discuss{$h$ denotes the element size} and on exterior facets \mer{the average and jump operators take the one-sided values}. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_h \times Q_h$ such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_h, \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

\subsubsection{Finite element solution of the perivascular network equations}

We consider meshes $\mathcal{I}_a, \mathcal{I}_v$ representing a
conforming subdivision of each of the perivascular networks
$\Lambda_a$, $\Lambda_v$. Relative to each mesh $\mathcal{I}$, we define the
space of (discontinuous) piecewise constants $\hat{Z}_h(\mathcal{I})$ and define
spaces of continuous piecewise linears $\hat{V}_{h, g}(\mathcal{I})$ with
prescribed boundary node values (on $\mathcal{N}$) given by $g$:
\begin{align}
  \hat{Z}_h(\mathcal{I}) &= \{z \in L^2(\Lambda) \, | \, z \vert_{\Lambda_i} \in \mathrm{P}^{0}(\Lambda_i) \, \foralls \Lambda^i \in \mathcal{I}\}, \\
  \hat{V}_{h, g}(\mathcal{I}) &= \{ v \in C^0(\Lambda) \, | \, v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda_i), \, v(x) = g(x) \, \foralls x \in \mathcal{N} \} .
  %Q_h^{\rm PVS} &= \{ q_h \in C^0(\Lambda),  q_h \vert_{\Lambda^i} \in \mathrm{P}^k(\Lambda_i), q_h \vert_{\Gamma_{\rm Pia}} = p_{\rm CSF} \}
\end{align}
To discretize~\eqref{eq:1d_flow} with the bifurcation
conditions~\eqref{eq:1d_flow:bifurcations} and boundary
conditions~\eqref{eq:1d_flow:bcs}, we use the space of continuous
functions $\hat{Z}_h(\mathcal{I})$ to enforce the continuity of the pressure at
bifurcation points, while the conservation of flux is enforced
(weakly) through the variational formulation. The discrete variational form of the equations then reads: Find $(\hat{q}, \hat{p}) \in \hat{Z}_h(\mathcal{I}) \times \hat{V}_{h, p_0}(\mathcal{I})$ such that
\begin{subequations}
\begin{align}
  \langle \hat{q}, z \rangle +  \langle z, \frac{A \kappa}{\mu} \partial_s \hat{p}  \rangle &= 0 \quad \forall z \in \hat{Z}_h(\mathcal{I}) \\
  \langle \hat{q}, \partial_s w \rangle &= 0 \quad  \forall w \in \hat{V}_{h, 0}(\mathcal{I}).
\end{align}
\end{subequations}%
where $\langle \cdot, \cdot \rangle$ denotes the $L^2(\mathcal{I})$-inner product and defined segment-wise. For a stability and convergence analysis of this discrete model, we refer to the reference\cite{gjerde2024directional}.

@@

@@

\mer{MER: This needs a pass to update to new notation and new style.}
